%
% Motivation / Introduction
%
\section{Introduction}

Operating systems (OSes) are specialised, low level software that manages a computer's
resources and provides a common interface for programs. The most common operating
systems used by most consumers today include MacOS and Windows. Xv6 is an operating
system written by MIT for the purposes of teaching [\iffalse TODO: add ref \fi].
As it is meant to be a teaching OS, it has very barebones functionality,
which includes a command line shell and some basic user programs by default.


%
% Aims, Objectives and deliverables.
%
\section{Aims, Objectives and Deliverables}

\subsection{Project Aims}

The aim of this project is to implement graphics support for Xv6. This includes
adding: basic PCIe support and driver for the graphics card, a simple interface
for user programs to draw to the screen, and possibly a grid window manager
to make use of Xv6's multicore CPU support.

\subsection{Objectives}

\begin{enumerate}
    \item{Implement basic PCIe support to detect and setup the QEMU standard VGA card}
    \item{Implement a driver for the VGA card to access registers and the framebuffer}
    \item{Implement a library to abstract away the intricacies of drawing to the screen and provide a basic interface to set pixels and draw shape primitives}
          \begin{enumerate}
              \item This is the bulk of the project, and many design decisions will have to be made
                    \begin{enumerate}
                        \item The main example is deciding whether or not the majority of the library will exist in userspace or kernelspace.

                              This will affect many things in development. Developing it in kernelspace will be easier as
                              there will be direct access to the framebuffer, but would result in a higher security risk
                              as any bugs and mistakes would have access to the kernel memory space. There would also be a
                              significantly large increase in system calls
                        \item Planning will be required for 'futureproofing' the library. For example there is an extended objective
                              to implement a window manager. Careful decisions will be required to ensure the graphics library can
                              interface easily with the windows.
                    \end{enumerate}
              \item Basic functionality like drawing shape primitive and applying simple 2D transformations
                    should be implemented. To support this, a maths utility library for dealing with the
                    graphics calculations should probably be implemented
          \end{enumerate}
    \item{(optional) Implement a grid window manager so multiple graphical user applications can run simultaneously}
\end{enumerate}

\subsection{Deliverables}

\begin{itemize}
    \item{Git repository containing source code for:}
          \begin{itemize}
              \item{Kernel PCIe configuration}
              \item{VGA driver}
              \item{Graphics library}
              \item{(optional) Grid window manager}
              \item{Example user programs}
          \end{itemize}
    \item{Demo video showcasing graphics implementation}
    \item{Final Report}
\end{itemize}


%
% Plan
%
\pagebreak
\section{Project Timeline}

Below is a gantt chart which splits the project into milestones with timeframe estimates. \\

\begin{ganttchart}[
        vgrid,
        x unit=0.4cm,
        milestone label font=\tiny,
        title label font=\tiny,
    ]{1}{30}
    \gantttitle{2021}{11}
    \gantttitle{2022}{19} \\
    \gantttitlelist{1,...,30}{1} \\
    \ganttgroup{Planning}{1}{8} \\
    \ganttgroup{Code Implementation}{8}{22} \\
    \ganttgroup{Report Finalisation}{22}{30} \\

    \ganttbar {Project Outline}{1}{6} \\
    \ganttbar {Preliminary Research}{4}{8} \\
    \ganttbar {Prototyping}{6}{8} \\

    \ganttbar {Intermediate Report}{8}{11} \\

    \ganttbar {PCIe Config}{8}{10} \\
    \ganttlinkedbar {VGA I/O}{10}{12} \\
    \ganttlinkedbar {VGA Config}{12}{14} \\
    \ganttlinkedbar {User Graphics Lib}{14}{18} \\
    \ganttlinkedbar {Example Programs}{18}{21}
    \ganttbar {}{24}{26} \\
    \ganttbar {Grid Window Manager}{21}{24} \\
    \ganttlink{elem11}{elem13}
    \ganttlink{elem13}{elem12}

    \ganttbar {Final Report}{14}{30} \\
\end{ganttchart}
\pagebreak


%
% Risk mitigation.
%
\section{Risk Mitigation}

Risks for this project are minor. Unexpected delays when implementing some of the
planned features are the only real risk, which is managed by splitting the project
into milestones with liberal estimates on the timeframes. If an issue arises that
severely impacts the time needed to complete the project, certain features can be
excluded such as the grid layout window manager.

For a better solution, the graphics library should be implemented in userspace.
This however increases the complexity of the project, so extra attention to time
will be required. If a feature takes too long, it will either be simplified or dropped.
If it is a critical feature, other planned less essential features will be simplified or
removed.

%
% Ethics.
%
\section{Legal \& Ethical issues}

Xv6 is made open source with the MIT License, which states any person may
\begin{quote}
    "..use, copy, \textbf{modify}, merge, publish,
    distribute, sublicense, and/or sell copies of the Software..."
\end{quote}
free of charge, given that the full license text and copyright notice
is provided.

As the software is being modified, to comply with the provided license the
original copyright and license will be included in the modified software.
Additionally, there are no references to the original authors' rights to intellectual
property of Xv6 derivatives, so the project - which are original modifications
made to Xv6 - will belong to the author (myself) and should be fine for submission.
% TODOS:

% license
% copyright & IP
